---
title: "ecocopula"
author: "Douglas Yu"
date: "04/10/2020"
output: html_document
---

```{r setup, eval=TRUE, include=TRUE}
# copy the following into each script   
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	include = FALSE
)

# script-specific libraries
suppressPackageStartupMessages({
  library(stringdist)
  library(vegan)
  library(mvabund)
  library(ordinal)
  library(ecoCopula)
  library(corrplot)
  library(ggrepel)
  library(RColorBrewer)
  library(tidygraph)
  library(ggraph)
})

# general-use packages
suppressPackageStartupMessages({
  library(here)
  library(fs)
  library(glue)
  library(tidyverse) # loads a bunch of packages (see below)
  library(readxl)
  library(cowplot)
  library(lubridate)
  library(patchwork)
  library(broom)
  library(ggeffects)
  library(viridis)
  library(arsenal) # for summary(comparedf())
  library(sjmisc) # for rotate_df()
  library(envDocument)
  library(inspectdf)
  library(conflicted)
})

conflict_prefer("filter", "dplyr", quiet = TRUE)
conflict_prefer("mutate", "dplyr", quiet = TRUE)
conflict_prefer("select", "dplyr", quiet = TRUE)
conflict_prefer("summarise", "dplyr", quiet = TRUE)
conflict_prefer("first", "dplyr", quiet = TRUE)
conflict_prefer("here", "here", quiet = TRUE)
conflict_prefer("separate", "tidyr", quiet = TRUE)
conflict_prefer("unite", "tidyr", quiet = TRUE)
conflict_prefer("intersect", "dplyr", quiet = TRUE)
conflict_prefer("setdiff", "dplyr", quiet = TRUE) # w/out this, R crashes
conflict_prefer("to_factor", "sjmisc", quiet = TRUE)
conflict_prefer("trim", "glue", quiet = TRUE)

# Provide real numbers, not scientific notation.
options(scipen = 999)
```

Ailaoshan_environmental.rdata obviously contains the environmental metadata

```{r working directory}
here() # should be: "/Users/Negorashi2011/Dropbox/Working_docs/Luo_Mingjie_Oregon/HJA_analyses_Kelpie/HJA_scripts/09_ecological_analyses"
```


```{r read data file}
samtoolsfilter <- "F2308" # F2308 filter only
samtoolsqual <- "q48"
minimaprundate <- 20200929
kelpierundate <- 20200927
primer <- "BF3BR2"
trap <- "M1"
period <- "S1"

(outputidxstatstabulatefolder <- glue("outputs_minimap2_{minimaprundate}_{samtoolsfilter}_{samtoolsqual}_kelpie{kelpierundate}_{primer}_vsearch97"))

(otuenvfilename <- glue("otuenv_{trap}{period}_minimap2_{minimaprundate}_kelpie{kelpierundate}.csv"))

otuenv <- read_csv(here("..", "..", "Kelpie_maps", 
                        outputidxstatstabulatefolder, 
                        otuenvfilename
                        )
                   )
```

```{r edit XY.csv}
# scale XY data
XY.csv <- otuenv %>% 
  select(UTM_E, UTM_N)
```

```{r edit otu.data}
# otu.data
# keep OTUs with >=5 incidences
minocc <- 3 # set to high number (e.g. 20) for testing
otu.qp.csv <- otuenv %>% select(contains("__"))
otu.qp.csv <- otu.qp.csv[ , specnumber(otu.qp.csv, MARGIN = 2) >= minocc] 

# convert to 0/1 data
otu.pa.csv <- otu.qp.csv
otu.pa.csv[otu.pa.csv > 0] <- 1
```

```{r edit env.scale.csv}
env.scale.csv <- otuenv %>% select(!contains("__")) %>% 
  select(-SiteName, -trap, -period, -UTM_E, -UTM_N, -clearcut) %>% 
  mutate(insideHJA = ifelse(insideHJA == "yes", 1, 0))

# # GIS + MS + LiDAR minimal set: gismslidarmin
env.scale.csv <- env.scale.csv %>%
  select(insideHJA, oldGrowthIndex, elevation_m, canopyHeight_m, minT_annual, distToRoad_m, distToStream_m, YrsSinceDist, mean.NDVI, mean.green, l_Cover_2m_4m, l_Cover_2m_max, l_Cover_4m_16m, l_p25, l_p95, l_rumple)

corrplot(cor(env.scale.csv), method = "ellipse", type = "lower")
```


```{r mvabund ecoCopula SSU}
source(here("1.1_Manyany.R"))

otu.qp.ord <- otu.qp.csv %>% 
  mutate(across(everything(), factor))

otu.qp.mod0 <- Manyany("clm", otu.qp.ord, abund ~ 1, data = env.scale.csv)

plot(otu.qp.mod0)

otu.qp.mod0.ord <- cord(otu.qp.mod0)
# plot(SSU_mod0_ord, biplot = TRUE)
```

```{r extract factors SSU}
site_res <- data.frame(otu.qp.mod0.ord$scores, env.scale.csv)
sp_res <- data.frame(otu.qp.mod0.ord$loadings, 
                     species = colnames(otu.qp.csv))
```

```{r ggplot ecoCopula SSU}
corrplot(cor(env.scale.csv), method = "ellipse", type = "lower")

alpha <- 4
ggplot() +
  geom_segment(aes(x = 0, y = 0, xend = Factor1 * alpha * 0.95, yend = Factor2 * alpha * 0.95), data = sp_res, size = .1) +
  geom_point(aes(x = Factor1, y = Factor2, color = elevation_m, size = YrsSinceDist), data = site_res) +
  geom_text_repel(aes(x = Factor1, y = Factor2, label = otuenv$SiteName), data = site_res, size = 2) +
  # geom_label_repel(aes(x = Factor1 * alpha, y = Factor2 * alpha, label = species), data = sp_res, size = 2.5) +
  scale_color_gradientn(colours = brewer.pal(n = 10, name = "PuOr")) +
  theme_classic() + # or "PuOr" or "RdYlBu"
  labs(title = "HJA, Manyany, ecoCopula")
```


From the Gordana Popovic tutorial:  "We can do this using the `mvabund` package, with function `manyany` as well as `clm` from the `ordinal` package. Sadly, R version 4.0.0 has broken `manyany`, and it has not been fixed yet, so I have a substitute called `Manyany` in the `1.1_Manyany.R` file which we will need to sourced in the code below. "

Manyany<-function (fn, yMat, formula, data, family = "negative.binomial", 
          composition = FALSE, block = NULL, get.what = "details", 
          var.power = NA, na.action = "na.exclude", ...) 
          

<details><summary>Reproducibility receipt</summary>

```{r}
# datetime
Sys.time()

# repository
git2r::repository(here::here())

envDocument::env_doc("table", git = FALSE)
```
