---
title: "ecocopula"
author: "Douglas Yu"
date: "04/10/2020"
output: html_document
---

```{r setup, eval=TRUE, include=TRUE}
# copy the following into each script   
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	include = FALSE
)

# script-specific libraries
suppressPackageStartupMessages({
  library(stringdist)
  library(vegan)
  library(mvabund)
  library(ordinal)
  library(ecoCopula)
  library(corrplot)
  library(ggrepel)
  library(RColorBrewer)
  library(tidygraph)
  library(ggraph)
})

# general-use packages
suppressPackageStartupMessages({
  library(here)
  library(fs)
  library(glue)
  library(tidyverse)
  library(readxl)
  library(cowplot)
  library(lubridate)
  library(patchwork)
  library(broom)
  library(ggeffects)
  library(viridis)
  library(arsenal) # for summary(comparedf())
  library(sjmisc) # for rotate_df()
  library(envDocument)
  library(inspectdf)
  library(conflicted)
})

conflict_prefer("filter", "dplyr", quiet = TRUE)
conflict_prefer("mutate", "dplyr", quiet = TRUE)
conflict_prefer("select", "dplyr", quiet = TRUE)
conflict_prefer("summarise", "dplyr", quiet = TRUE)
conflict_prefer("first", "dplyr", quiet = TRUE)
conflict_prefer("here", "here", quiet = TRUE)
conflict_prefer("separate", "tidyr", quiet = TRUE)
conflict_prefer("unite", "tidyr", quiet = TRUE)
conflict_prefer("intersect", "dplyr", quiet = TRUE)
conflict_prefer("setdiff", "dplyr", quiet = TRUE) # w/out this, R crashes
conflict_prefer("to_factor", "sjmisc", quiet = TRUE)
conflict_prefer("trim", "glue", quiet = TRUE)

# Provide real numbers, not scientific notation.
options(scipen = 999)
```


```{r working directory}
here() # should be: "/Users/Negorashi2011/Dropbox/Working_docs/Luo_Mingjie_Oregon/HJA_analyses_Kelpie/HJA_scripts/09_ecological_analyses"
```


```{r read data file}
samtoolsfilter <- "F2308" # F2308 filter only
samtoolsqual <- "q48"
minimaprundate <- 20200929
kelpierundate <- 20200927
primer <- "BF3BR2"
trap <- "M1"
period <- "S1"

(outputidxstatstabulatefolder <- glue("outputs_minimap2_{minimaprundate}_{samtoolsfilter}_{samtoolsqual}_kelpie{kelpierundate}_{primer}_vsearch97"))

(otuenvfilename <- glue("otuenv_{trap}{period}_minimap2_{minimaprundate}_kelpie{kelpierundate}.csv"))

otuenv <- read_csv(here("..", "..", "Kelpie_maps", 
                        outputidxstatstabulatefolder, 
                        otuenvfilename
                        )
                   )
```

```{r edit XY.csv}
# scale XY data
XY.csv <- otuenv %>% 
    select(UTM_E, UTM_N) %>% 
    scale() %>% 
    as_tibble()
```

```{r edit otu.data}
# otu.data
# keep OTUs with >=5 incidences
minocc <- 5 # set to high number (e.g. 20) for testing
otu.qp.csv <- otuenv %>% select(contains("__"))
otu.qp.csv <- otu.qp.csv[ , specnumber(otu.qp.csv, MARGIN = 2) >= minocc] 

# convert to 0/1 data
otu.pa.csv <- otu.qp.csv
otu.pa.csv[otu.pa.csv > 0] <- 1
```

```{r edit env.scale.csv}
env.scale.csv <- otuenv %>% 
    select(!contains("__")) %>% 
    select(-SiteName, -trap, -period, -UTM_E, -UTM_N, 
           -clearcut, -oldGrowthIndex, -starts_with("nor.")) %>%
    mutate(insideHJA = ifelse(insideHJA == "yes", 1, 0))

# # GIS + MS + LiDAR:  gismslidar
msdate <- c("20180726") # choose date of multispectral data
env.scale.csv <- env.scale.csv %>%
  select(insideHJA, elevation_m, canopyHeight_m, minT_annual, precipitation_mm, distToRoad_m, distToStream_m, YrsSinceDist, contains(all_of(msdate)), l_Cover_2m_max, l_Cover_2m_4m, l_Cover_4m_16m, l_p25, l_rumple)

corrplot(cor(env.scale.csv), method = "ellipse", type = "lower")
```

Run either "Manyany qp ecoCopula" or "manyglm pa ecoCopula"
```{r ordinal stackedsdm qp ecoCopula}
abund <- "qp"
glue("otu.{abund}.csv")
get(glue("otu.{abund}.csv"))[1:5, 1:5]

# beta
otu.ord <- (get(glue("otu.{abund}.csv")))
otu.mod0 <- stackedsdm(otu.ord, formula_X = ~ 1, data = env.scale.csv,
                       family="beta", ncores = 4)

# ordinal
otu.ord <- (get(glue("otu.{abund}.csv"))) %>% 
    mutate(across(everything(), ~rank(.x, ties.method = "min")))
otu.mod0 <- stackedsdm(otu.ord, formula_X = ~ 1, data = env.scale.csv,
                       family="ordinal", ncores = 4)

plot(otu.mod0) # chk residuals

otu.mod0.ord <- cord(otu.mod0)
# plot(SSU_mod0_ord, biplot = TRUE)
```
OR
```{r Manyany qp ecoCopula}
source(here("1.1_Manyany.R"))

abund <- "qp"
glue("otu.{abund}.csv")

otu.ord <- get(glue("otu.{abund}.csv")) %>% 
  mutate(across(everything(), factor))

otu.mod0 <- Manyany("clm", otu.ord, abund ~ 1, data = env.scale.csv)

plot(otu.mod0) # chk residuals

otu.mod0.ord <- cord(otu.mod0)
# plot(SSU_mod0_ord, biplot = TRUE)
```
OR
```{r manyglm pa ecoCopula}
abund <- "pa"
glue("otu.{abund}.csv")

otu.ord <- mvabund(get(glue("otu.{abund}.csv")))
is.mvabund(otu.ord)

otu.mod0 <- manyglm(otu.ord ~ 1, family = "binomial")

plot(otu.mod0) # chk residuals

otu.mod0.ord <- cord(otu.mod0)
# plot(SSU_mod0_ord, biplot = TRUE)
```

Process output of cord()
```{r extract factors}
site_res <- data.frame(otu.mod0.ord$scores, env.scale.csv, XY.csv)

sp_res <- data.frame(otu.mod0.ord$loadings, 
                     species = colnames(get(glue("otu.{abund}.csv")))) %>%
    separate(species, into = c("OTU", "empty", "class", "order", "family",
                               "genus", "epithet", "BOLD", "BOLDID",
                               "size"), remove = FALSE, sep = "_") %>% 
    select(-empty)
```

```{r ggplot ecoCopula}
corrplot(cor(site_res), method = "ellipse", type = "lower", 
         title = glue({abund}))

# save corrplot
pdf(file = here("..", "..", "Kelpie_maps", 
                outputidxstatstabulatefolder,
                glue("corrplot_{trap}{period}_minimap2_{minimaprundate}_kelpie{kelpierundate}_{abund}.pdf")
                )
    )

corrplot(cor(site_res), method = "ellipse", type = "lower", 
         title = glue({abund}))
dev.off()
```

```{r plot ecoCopula ordination}
alpha <- 4
p0 <- ggplot() + 
  geom_segment(aes(x = 0, y = 0, 
                   xend = Factor1 * alpha * 0.95, 
                   yend = Factor2 * alpha * 0.95), 
               data = sp_res, 
               size = .1) +
  geom_point(aes(x = Factor1, y = Factor2, 
                 color = elevation_m, 
                 size = YrsSinceDist, 
                 shape = as.factor(insideHJA)
                 ), 
             data = site_res) + 
  geom_text_repel(aes(x = Factor1, y = Factor2, 
                      label = otuenv$SiteName), 
                  data = site_res, 
                  size = 2) +
  # geom_label_repel(aes(x = Factor1 * alpha, y = Factor2 * alpha, label = species), data = sp_res, size = 2.5) +
  scale_color_gradientn(colours = brewer.pal(n = 10, name = "RdYlBu")) +
  theme_classic() + # or "PuOr" or "RdYlBu"
  labs(title = glue("HJA, ecoCopula ordination, {abund}"))

p0
```

```{r plot by XY position}
p1 <- ggplot() + 
  geom_point(aes(x = UTM_E, y = UTM_N, 
                 color = Factor1, 
                 size = YrsSinceDist, # YrsSinceDist, l_rumple
                 shape = as.factor(insideHJA)
                 ), 
             data = site_res) +
  geom_text_repel(aes(x = UTM_E, y = UTM_N, 
                      label = otuenv$SiteName), 
                  data = site_res, 
                  size = 2) +
  # geom_label_repel(aes(x = Factor1 * alpha, y = Factor2 * alpha, label = species), data = sp_res, size = 2.5) +
  scale_color_gradientn(colours = brewer.pal(n = 10, name = "PuOr")) +
  theme_classic() + # or "PuOr" or "RdYlBu"
  labs(title = glue("HJA, XY plot, Factor1, {abund}"))

p2 <- ggplot() + 
  geom_point(aes(x = UTM_E, y = UTM_N, 
                 color = Factor2, 
                 size = YrsSinceDist, # YrsSinceDist, l_rumple
                 shape = as.factor(insideHJA)
                 ), 
             data = site_res) +
  geom_text_repel(aes(x = UTM_E, y = UTM_N, 
                      label = otuenv$SiteName), 
                  data = site_res, 
                  size = 2) +
  # geom_label_repel(aes(x = Factor1 * alpha, y = Factor2 * alpha, label = species), data = sp_res, size = 2.5) +
  scale_color_gradientn(colours = brewer.pal(n = 10, name = "PuOr")) +
  theme_classic() + # or "PuOr" or "RdYlBu"
  labs(title = glue("HJA, XY plot, Factor2, {abund}"))

p1 | p2
```

```{r save ecoCopula plots}
# pdf(width = 11.7, height = 8.3, file = here("..", "..", "Kelpie_maps", 
#                 outputidxstatstabulatefolder,
#                 glue("ecoCopula_{trap}{period}_minimap2_{minimaprundate}_kelpie{kelpierundate}_{abund}.pdf")
#                 )
#     )
# p0
# dev.off()

ggsave(filename = here("..", "..", "Kelpie_maps", 
                outputidxstatstabulatefolder,
                glue("ecoCopula_{trap}{period}_minimap2_{minimaprundate}_kelpie{kelpierundate}_{abund}.pdf")
                ), 
       width = 11.7, 
       height = 8.3, 
       device = "pdf",
       p0)

# pdf(width = 11.7, height = 8.3, file = here("..", "..", "Kelpie_maps", 
#                 outputidxstatstabulatefolder,
#                 glue("ecoCopulaXY_{trap}{period}_minimap2_{minimaprundate}_kelpie{kelpierundate}_{abund}.pdf")
#                 )
#     )
# p1 | p2
# dev.off()

ggsave(filename = here("..", "..", "Kelpie_maps", 
                outputidxstatstabulatefolder,
                glue("ecoCopulaXY_{trap}{period}_minimap2_{minimaprundate}_kelpie{kelpierundate}_{abund}.pdf")
                ), 
       width = 11.7, 
       height = 8.3, 
       device = "pdf",
       p1 | p2)
```

From the Gordana Popovic tutorial:  "We can do this using the `mvabund` package, with function `manyany` as well as `clm` from the `ordinal` package. Sadly, R version 4.0.0 has broken `manyany`, and it has not been fixed yet, so I have a substitute called `Manyany` in the `1.1_Manyany.R` file which we will need to sourced in the code below. "

Manyany<-function (fn, yMat, formula, data, family = "negative.binomial", 
          composition = FALSE, block = NULL, get.what = "details", 
          var.power = NA, na.action = "na.exclude", ...) 
          

<details>
<summary>Reproducibility receipt</summary>
```{r reproducibility receipt}
# datetime
Sys.time()

# repository
git2r::repository(here::here())

envDocument::env_doc("table", git = FALSE)
```
