---
title: "FSL_correction"
author: "Douglas Yu"
date: "02/10/2020"
output: html_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(lubridate)
library(knitr)
library(beepr)
library(arsenal) # for summary(comparedf())
library(sjmisc) # for rotate_df()
library(pivottabler)
library(vegan)
library(fs)
library(furrr)
library(glue)
library(here)
library(scales)
library(conflicted)
  conflict_prefer("mutate", "dplyr", quiet = TRUE)
  conflict_prefer("select", "dplyr", quiet = TRUE)
  conflict_prefer("summarise", "dplyr", quiet = TRUE)
  conflict_prefer("filter", "dplyr", quiet = TRUE)
  conflict_prefer("first", "dplyr", quiet = TRUE)
  conflict_prefer("here", "here", quiet = TRUE)
  conflict_prefer("separate", "tidyr", quiet = TRUE)
  conflict_prefer("unite", "tidyr", quiet = TRUE)
  conflict_prefer("trim", "sjmisc", quiet=TRUE)
  conflict_prefer("rescale", "scales", quiet=TRUE)

# Real numbers, not scientific notation.
options(scipen = 999)
```

```{r read species_by_sample table}
samtoolsfilter <- "F2308" # F2308 filter only
samtoolsqual <- "q48"
minimaprundate <- 20200929
kelpierundate <- 20200927
primer <- "BF3BR2"

outputidxstatstabulatefolder <- glue("outputs_minimap2_{minimaprundate}_{samtoolsfilter}_{samtoolsqual}_kelpie{kelpierundate}_{primer}_vsearch97")

idx_meta_genomecov_filtered_wide <- read_csv(here("..", "..", "Kelpie_maps", outputidxstatstabulatefolder, glue("sample_by_species_table_{samtoolsfilter}_minimap2_{minimaprundate}_kelpie{kelpierundate}.csv")))
```

read in lysis data, correct sample names, remove missing samples, separate into site, trap, period
```{r read FSL data}
lysis <- read_csv(here("..", 
                       "08_reference_sequences_datasets",
                       "lysis_buffer_volume_20200724.csv")
                  ) %>% 
  select(-"WEIGHT(g)") %>% 
  rename(lysisBufferml = "added lysis buffer(ml)") %>% 
  mutate(sample = str_replace(sample, "HOBO_", "HOBO-")) %>%
  mutate(sample = str_replace(sample, "SM_", "SM-")) %>% 
  mutate(sample = str_replace(sample, "HOBO-043_M1_S2", "HOBO-063_M1_S2")) %>%
  filter(sample != "SM-05_M1_S1") %>% 
  filter(sample != "HOBO-351_M1_S2") %>% 
  filter(sample != "HOBO-357_M1_S2") %>% 
  filter(sample != "HOBO-036_M1_S1") %>%
  filter(sample != "HOBO-036_M1_S2") %>%
  filter(sample != "268453_M1_S2") %>% 
  filter(sample != "HOBO-317_M1_S1") %>% 
  separate(sample, into = c("SiteName", "trap", "period"), sep = "_", remove = TRUE) 

# HOBO-063_M1_S2 was misnamed HOBO-043_M1_S2 in lysis_buffer_volume_20200724.csv

# from 1.0_idxstats_tabulate.Rmd
  # SM-05_M1_S1 had no spike reads and was thus removed from idx_meta_genomecov_filtered_wide
  # HOBO-357-M1-S2 and HOBO-351-M1-S2 were ambiguously named and were thus removed from idx_meta_genomecov_filtered_wide
  # HOBO-036-M1 lacked info on session (S1/S2) and thus was removed from idx_meta_genomecov_filtered_wide

# 268453_M1_S2 is an extra line in lysis_buffer_volume_20200724.csv, but not in OTU table
# HOBO-317_M1_S1 is an extra line in lysis_buffer_volume_20200724.csv, but not in OTU table
```

Join lysis and idx_meta_genomecov_filtered_wide
```{r}
idx_meta_genomecov_filtered_lysis <- lysis %>% 
  full_join(idx_meta_genomecov_filtered_wide, 
            by = c("SiteName" = "SiteName", "trap" = "trap", "period" = "period")
            ) %>% 
  filter(!(SiteName == "SM-06" & trap == "M1" & period == "S1")) %>% 
  relocate(SiteName, trap, period, lysisBufferml, aliquot, starts_with("COI-SPIKE")) %>% 
  rename(COISpikeElaterid = `COI-SPIKE__Coleoptera_Elateridae_0.8_NA_NA_NA_NA_NA`,
         COISpikeMordellid = `COI-SPIKE__Coleoptera_Mordellidae_0.4_NA_NA_NA_NA_NA`) # backticks because the "-" causes problems
# SM-06-M1_S1 is in OTU table but missing from lysis_buffer_volume_20200724.csv, or maybe this is the missing SM-05_M1_S1 (but i know that i removed SM-05_M1_S1 from the OTU table because it got no spike reads)
```

To be checked with Mingjie:

1. 268453_M1_S2 is in lysis_buffer_volume_20200724.csv, but not in OTU table. 
2. HOBO-317_M1_S1 is in lysis_buffer_volume_20200724.csv, but not in OTU table
3. SM-06-M1_S1 is missing from lysis_buffer_volume_20200724.csv, but is in OTU table
4. HOBO-063_M1_S2 is missing from lysis_buffer_volume_20200724.csv but i think that HOBO-043_M1_S2 is this sample (there is no HOBO-043 site)


spike-reads and lysis buffer correction factors

```{r}
idx_meta_genomecov_filtered_lysis <- idx_meta_genomecov_filtered_lysis %>% 
  mutate(spike_wt_mean = (COISpikeElaterid/9 + COISpikeMordellid)/2) %>% 
  relocate(spike_wt_mean, .after = COISpikeMordellid)

hist(idx_meta_genomecov_filtered_lysis$spike_wt_mean)

idx_meta_genomecov_filtered_lysis <- idx_meta_genomecov_filtered_lysis %>% 
  mutate(lysis_ratio = aliquot / lysisBufferml) %>% 
  relocate(lysis_ratio, .after = aliquot)

hist(idx_meta_genomecov_filtered_lysis$lysis_ratio)
```

_FSL corrections_ from Ji et al. (2020) SPIKEPIPE paper
1. Lysis buffer correction. – For each sample (row), calculate lysis_ratio, which is the proportion of the total lysis buffer that was used for DNA extraction. For example, if half the lysis buffer of a sample was used, lysis_ratio is 0.5. Finally, divide the read number of each species in the row by lysis_ratio 

2. Spike-in correction. – (a) For each sample (row), calculate the spike_wt_mean. Divide COISpikeElaterid by 9 and then take the mean of COISpikeElaterid and COISpikeMordellid. Finally, divide the read number of each species in the row by spike_wt_mean. 

_Quasi-probability correction_
3. For each species (column), scale to a quasiprobability: scales::rescale(log(x + 0.001))

Using mutate(across()) to apply the 3 corrections
```{r correct read numbers}
idx_meta_genomecov_filtered_lysis <- idx_meta_genomecov_filtered_lysis %>% 
  mutate(across(contains("__"), ~ .x /spike_wt_mean)) %>% 
  mutate(across(contains("__"), ~ .x /lysis_ratio)) %>% 
  mutate(across(contains("__"), ~ rescale(log(.x + 0.001)))) # {rescale}

hist(idx_meta_genomecov_filtered_lysis$`R1240-70__Insecta_Diptera_Syrphidae_Blera_scitula_BOLD_ABY7981_size=998`)
```

select columns and scale environmental variables
```{r}
idx_meta_genomecov_filtered_lysis <- idx_meta_genomecov_filtered_lysis %>% 
  select(-lysisBufferml, -aliquot, -lysis_ratio, -COISpikeElaterid,
         -COISpikeMordellid, -spike_wt_mean, -`lysis batch`, -site_trap_period) %>%
  relocate(insideHJA, oldGrowthIndex, .after = clearcut)

idx_meta_genomecov_filtered_lysis <- idx_meta_genomecov_filtered_lysis %>% 
  mutate(across(oldGrowthIndex:l_rumple, scale))
```



select subsets by trap and period, remove OTUs with column sums ==0 (OTUs that are not present in the subdataset)
```{r}
otuenv_M1S1 <- idx_meta_genomecov_filtered_lysis %>% 
  filter(trap == "M1" & period == "S1") %>% 
  select(which(map_lgl(., 
                       ~contains("__") && 
                         specnumber(.x, MARGIN=2) > 0) # {vegan}
               )
         ) 

otuenv_M2S1 <- idx_meta_genomecov_filtered_lysis %>% 
  filter(trap == "M2" & period == "S1") %>% 
  select(which(map_lgl(., 
                       ~contains("__") && 
                         specnumber(.x, MARGIN=2) > 0) # {vegan}
               )
         ) 

otuenv_M1S2 <- idx_meta_genomecov_filtered_lysis %>% 
  filter(trap == "M1" & period == "S2") %>% 
  select(which(map_lgl(., 
                       ~contains("__") && 
                         specnumber(.x, MARGIN=2) > 0) # {vegan}
               )
         ) 

otuenv_M2S2 <- idx_meta_genomecov_filtered_lysis %>% 
  filter(trap == "M2" & period == "S2") %>% 
  select(which(map_lgl(., 
                       ~contains("__") && 
                         specnumber(.x, MARGIN=2) > 0) # {vegan}
               )
         ) 
```



```{r}
outputidxstatstabulatefolder # confirm folder to save outputs in
minimaprundate # confirm minimaprundate (same as in outputidxstatstabulatefolder)
kelpierundate # confirm kelpierundate (same as in outputidxstatstabulatefolder)

writefile <- function(trap, period) {
  otuenv <- glue("otuenv_{trap}{period}")
  write_csv(get(otuenv), 
            here("..", "..", "Kelpie_maps", 
                 outputidxstatstabulatefolder,
    glue("otuenv_{trap}{period}_minimap2_{minimaprundate}_kelpie{kelpierundate}.csv")))
}

writefile(trap="M1", period="S1")
writefile("M2", "S1")
writefile("M1", "S2")
writefile("M2", "S2")
```



<details><summary>Reproducibility receipt</summary>
```{r}
# datetime
Sys.time() 

# repository 
git2r::repository(here::here()) 

# environment doc
envDocument::env_doc("table", git = FALSE) 
```
