---
title: "dataprep"
author: "Douglas Yu"
date: "30/08/2020"
output: html_document
---

Prepares datasets for sjSDM analysis

1.  Reads in one of the otuenv sub-datasets output by 2.1_FSL_correction.Rmd (currently M1S1). This file has been log(FSL)-corrected, subset to M1S1, M2S1, M1S2, or M2S2, and saved as a csv file:
    otuenv_M1S1_minimap2_20200929_kelpie20200927.csv
    otuenv_M2S1_minimap2_20200929_kelpie20200927.csv
    otuenv_M1S2_minimap2_20200929_kelpie20200927.csv
    otuenv_M2S2_minimap2_20200929_kelpie20200927.csv

2.  Creates a scaled XY file for sampling locations

3.  Creates two sample X OTU tables, limited to OTUs that appear >=minocc times (currently 5), one with quasiprobabilities and one with presence-absence:  otu.pa.csv and otu.qp.csv

4.  Create an environmental covariates table from a subset of all the env covariates, some of which are averaged, some are logged, and numeric ones are scaled

5.  Save the files to a data folder in the sjSDM folder

```{r setup, eval=TRUE, include=TRUE}
# script-specific libraries
suppressPackageStartupMessages({
  library(corrplot)
  library(vegan)
})

# general-use packages
suppressPackageStartupMessages({
  library(here)
  library(fs)
  library(glue)
  library(tidyverse) # loads a bunch of packages (see below)
  library(readxl)
  library(cowplot)
  library(lubridate)
  library(patchwork)
  library(broom)
  library(ggeffects)
  library(viridis)
  library(arsenal) # for summary(comparedf())
  library(waldo) # for compare()
  library(sjmisc) # for rotate_df()
  library(envDocument)
  library(inspectdf)
  library(conflicted)
})

conflict_prefer("filter", "dplyr", quiet = TRUE)
conflict_prefer("mutate", "dplyr", quiet = TRUE)
conflict_prefer("select", "dplyr", quiet = TRUE)
conflict_prefer("summarise", "dplyr", quiet = TRUE)
conflict_prefer("first", "dplyr", quiet = TRUE)
conflict_prefer("here", "here", quiet = TRUE)
conflict_prefer("separate", "tidyr", quiet = TRUE)
conflict_prefer("unite", "tidyr", quiet = TRUE)
conflict_prefer("intersect", "dplyr", quiet = TRUE)
conflict_prefer("setdiff", "dplyr", quiet = TRUE) # w/out this, R crashes
conflict_prefer("to_factor", "sjmisc", quiet = TRUE)
conflict_prefer("trim", "glue", quiet = TRUE)

# Provide real numbers, not scientific notation.
options(scipen = 999)
```

```{r working directory}
here() # should be: "/Users/Negorashi2011/Dropbox/Working_docs/Luo_Mingjie_Oregon/HJA_analyses_Kelpie/HJA_scripts/07_idxstats_tabulate"
```

```{r read otuenv file}
samtoolsfilter <- "F2308" # F2308 filter only
samtoolsqual <- "q48"
minimaprundate <- 20200929
kelpierundate <- 20200927
primer <- "BF3BR2"
trap <- "M1"
period <- "S1"

(outputidxstatstabulatefolder <- glue("outputs_minimap2_{minimaprundate}_{samtoolsfilter}_{samtoolsqual}_kelpie{kelpierundate}_{primer}_vsearch97"))

(otuenvfilename <- glue("otuenv_{trap}{period}_minimap2_{minimaprundate}_kelpie{kelpierundate}.csv"))

otuenv <- read_csv(here("..", "..", "Kelpie_maps", 
                        outputidxstatstabulatefolder, 
                        otuenvfilename
                        )
                   )
```

```{r create XY.csv}
# scale XY data
XY.csv <- otuenv %>% 
  select(UTM_E, UTM_N) %>% 
  scale() %>% 
  as_tibble()
```

```{r create otu.pa.csv and otu.qp.csv}
# otu.data
# keep OTUs with >=5 incidences
minocc <- 5 # set to high number (e.g. 20) for testing
otu.qp.csv <- otuenv %>% select(contains("__"))
otu.qp.csv <- otu.qp.csv[ , specnumber(otu.qp.csv, MARGIN = 2) >= minocc] 

# convert to 0/1 data
otu.pa.csv <- otu.qp.csv
otu.pa.csv[otu.pa.csv > 0] <- 1

min(colSums(otu.pa.csv)) == minocc # should be TRUE

# otu.pa.csv_test <- vegan::decostand(otu.qp.csv, method = "pa")
# waldo::compare(otu.pa.csv, otu.pa.csv_test)
# summcomparedf <- summary(comparedf(otu.pa.csv, otu.pa.csv_test))
```

log and scale env variables
```{r create scale.env.csv}
# default is all environmental covariates: GIS + MS + Lidar

scale.env.csv <- otuenv %>% 
  select(!contains("__")) %>% # remove OTUs
  select(-SiteName, -trap, -period, -UTM_E, -UTM_N, -lysis_ratio, 
         -COISpike_sum, -starts_with("nor")) # %>% 
  # mutate(insideHJA = ifelse(insideHJA == "yes", 1, 0)) %>% 
  # mutate(clearcut = ifelse(clearcut == "yes", 1, 0))

names(scale.env.csv)
#  [1] "clearcut"           "insideHJA"          "oldGrowthIndex"    
#  [4] "elevation_m"        "canopyHeight_m"     "minT_annual"       
#  [7] "maxT_annual"        "precipitation_mm"   "distToRoad_m"      
# [10] "distToStream_m"     "YrsSinceDist"       "B1_20180717"       
# [13] "B2_20180717"        "B3_20180717"        "B4_20180717"       
# [16] "B5_20180717"        "B6_20180717"        "B7_20180717"       
# [19] "B10_20180717"       "B11_20180717"       "NDVI_20180717"     
# [22] "EVI_20180717"       "B_20180717"         "G_20180717"        
# [25] "W_20180717"         "B1_20180726"        "B2_20180726"       
# [28] "B3_20180726"        "B4_20180726"        "B5_20180726"       
# [31] "B6_20180726"        "B7_20180726"        "B10_20180726"      
# [34] "B11_20180726"       "NDVI_20180726"      "EVI_20180726"      
# [37] "B_20180726"         "G_20180726"         "W_20180726"        
# [40] "B1_20180802"        "B2_20180802"        "B3_20180802"       
# [43] "B4_20180802"        "B5_20180802"        "B6_20180802"       
# [46] "B7_20180802"        "B10_20180802"       "B11_20180802"      
# [49] "NDVI_20180802"      "EVI_20180802"       "B_20180802"        
# [52] "G_20180802"         "W_20180802"         "B1_20180818"       
# [55] "B2_20180818"        "B3_20180818"        "B4_20180818"       
# [58] "B5_20180818"        "B6_20180818"        "B7_20180818"       
# [61] "B10_20180818"       "B11_20180818"       "NDVI_20180818"     
# [64] "EVI_20180818"       "B_20180818"         "G_20180818"        
# [67] "W_20180818"         "mean.NDVI"          "mean.EVI"          
# [70] "mean.bright"        "mean.green"         "mean.wet"          
# [73] "mean.NDVI.scale"    "mean.EVI.scale"     "mean.green.scale"  
# [76] "mean.bright.scale"  "mean.wet.scale"     "l_Cover_2m_max"    
# [79] "l_Cover_2m_max_all" "l_Cover_2m_4m"      "l_Cover_2m_4m_all" 
# [82] "l_Cover_4m_16m"     "l_p25"              "l_p25_all"         
# [85] "l_p95"              "l_p95_all"          "l_rumple"         

corrplot(cor(scale.env.csv[, -c(1:2)]), method = "ellipse", type = "lower", tl.cex = 0.5)

# GIS + MS + LiDAR:  gismslidar
# average, optionally log, select, and scale env covariates
# using a set of covariates chosen by Christian
scale.env.csv <- scale.env.csv %>%
    mutate(B1_mean = rowMeans(across(starts_with("B1_")))) %>% 
    mutate(B4_mean = rowMeans(across(starts_with("B4_")))) %>%  
    mutate(lg_DistStream = log(distToStream_m + 0.001)) %>% 
    mutate(lg_DistRoad = log(distToRoad_m + 0.001)) %>% 
    mutate(lg_YrsDisturb = log(YrsSinceDist + 0.001)) %>% 
    mutate(lg_cover2m_max = log(l_Cover_2m_max + 0.001)) %>% 
    mutate(lg_cover2m_4m = log(l_Cover_2m_4m + 0.001)) %>%
    mutate(lg_cover4m_16m = log(l_Cover_4m_16m + 0.001)) %>%   
    select(clearcut, insideHJA, oldGrowthIndex, elevation_m, canopyHeight_m, precipitation_mm, mean.NDVI, mean.EVI, mean.green, mean.wet, l_p25, l_rumple, B1_mean, B4_mean, lg_DistStream, lg_DistRoad, lg_YrsDisturb, lg_cover2m_max, lg_cover2m_4m, lg_cover4m_16m) %>% 
    mutate(across(where(is.numeric), scale))

# msdate <- c("20180717") # alternative "20180726" # date of MS data 
# logenv <- c("distToRoad_m", "distToStream_m", "YrsSinceDist") # 3 variables for which the short distances are more important than long distances
# scale.env.csv <- scale.env.csv %>%
#   select(insideHJA, elevation_m, canopyHeight_m, minT_annual, precipitation_mm, distToRoad_m, distToStream_m, YrsSinceDist, contains(all_of(msdate)), starts_with("mean."), l_Cover_2m_max, l_Cover_2m_4m, l_Cover_4m_16m, l_p25, l_p95, l_rumple) %>% 
#     select(!ends_with(".scale")) %>% 
#     mutate(across(all_of(logenv), ~log(.x + 0.001))) %>% # optionally comment out
#     mutate(across(everything(), scale))

corrplot(cor(scale.env.csv[, -c(1:2)]), method = "ellipse", type = "lower", tl.cex = 0.5)
```


These files are used as input to the sjsdm_cv() and sjsdm() code

```{r save data files}
# set variables
prepdate <- 20201213 # data prep date
minocc <- 5 # minimum occupancy (incidence) per OTU
envvar <- "gismslidar" # gismslidarmin, gismslidar, gis, ms, lidar, mslidar
(datafolder <- glue("data_{prepdate}_{minocc}minocc_{envvar}"))

dir_create(here("..", "..", "Kelpie_maps", outputidxstatstabulatefolder, "crossvalidation_data", datafolder))

write_csv(scale.env.csv, here("..", "..", "Kelpie_maps", outputidxstatstabulatefolder, "crossvalidation_data", datafolder, "scale.env.csv"))

write_csv(XY.csv, here("..", "..", "Kelpie_maps", outputidxstatstabulatefolder, "crossvalidation_data", datafolder, "XY.csv"))

write_csv(otu.qp.csv, here("..", "..", "Kelpie_maps", outputidxstatstabulatefolder, "crossvalidation_data", datafolder,  "otu.qp.csv"))

write_csv(otu.pa.csv, here("..", "..", "Kelpie_maps", outputidxstatstabulatefolder, "crossvalidation_data", datafolder, "otu.pa.csv"))
```

Upload the datafiles to ~/Hja_sjsdm/

<details><summary>Reproducibility receipt</summary>
```{r}
# datetime
Sys.time()

# repository
git2r::repository(here::here())

env_doc("table", git = FALSE)
```
</details>
