---
title: "Process additional sjsdm output"
author: "Douglas Yu"
date: "07/12/2020"
output: html_document
---

Produce additional sjSDM outputs:
Correlation heatmap
Circular covariance/correlation plot sorted by abundance
Circular covariance/correlation plot sorted by env covariates
Env coefficients by species barchart

run on the laptop

```{r setup, eval=TRUE, include=TRUE}
# script-specific libraries
suppressPackageStartupMessages({
    library(stringdist)
    library(vegan)
    library(gridBase)
    library(grid)
    library(gridExtra)
    library(ggcorrplot)
    library(reticulate)
    library(sjSDM)
    library(mgsub)
    library(vcd)
    library(RColorBrewer)
    library(reshape2)
})

# general-use packages
suppressPackageStartupMessages({
  library(tidyverse) # loads a bunch of packages (see below)
  library(here)
  library(fs)
  library(glue)
  library(readxl)
  library(cowplot)
  library(lubridate)
  library(patchwork)
  library(broom)
  library(ggeffects)
  library(viridis)
  library(arsenal) # for tableby()
  library(waldo) # for compare()
  library(sjmisc) # for rotate_df()
  library(envDocument)
  library(inspectdf)
  library(conflicted)
  library(knitr)
  library(beepr)
  library(pivottabler)
  library(furrr)
  library(scales)
  library(janitor)
})

conflict_prefer("mutate", "dplyr", quiet = TRUE)
conflict_prefer("select", "dplyr", quiet = TRUE)
conflict_prefer("summarise", "dplyr", quiet = TRUE)
conflict_prefer("filter", "dplyr", quiet = TRUE)
conflict_prefer("first", "dplyr", quiet = TRUE)
conflict_prefer("here", "here", quiet = TRUE)
conflict_prefer("separate", "tidyr", quiet = TRUE)
conflict_prefer("unite", "tidyr", quiet = TRUE)
conflict_prefer("trim", "sjmisc", quiet=TRUE)
conflict_prefer("rescale", "scales", quiet=TRUE)
conflict_prefer("rescale", "discard", quiet=TRUE)
conflict_prefer("intersect", "dplyr", quiet = TRUE)
conflict_prefer("setdiff", "dplyr", quiet = TRUE) # w/out this, R crashes
conflict_prefer("to_factor", "sjmisc", quiet = TRUE)
conflict_prefer("trim", "glue", quiet = TRUE)
conflict_prefer("discard", "purrr", quiet = TRUE)
conflict_prefer("importance", "sjSDM", quiet = TRUE)
# Real numbers, not scientific notation.
options(scipen = 999)
```

```{r load functions}
packageVersion('sjSDM')
# [1] ‘0.1.2.9000’
	
source(here("source", "sjsdm_function.r"))
source(here("source", "sjsdm-analyse-functions.r"))
```

samtoolsfilter <- "F2308" # F2308 filter only
samtoolsqual <- "q48"
minimaprundate <- 20200929
kelpierundate <- 20200927
primer <- "BF3BR2"
trap <- "M1"
period <- "S1"

(outputidxstatstabulatefolder <- glue("outputs_minimap2_{minimaprundate}_{samtoolsfilter}_{samtoolsqual}_kelpie{kelpierundate}_{primer}_vsearch97"))


```{r set names}
# load data from 'crossvalidation' folder 
samtoolsfilter <- "F2308" # F2308 filter only
samtoolsqual <- "q48"
minimaprundate <- 20200929
kelpierundate <- 20200927
primer <- "BF3BR2"

date.cross.validation = 20201119
minocc = 5
envvar = "gismslidar"
abund = "qp" # "qp" or "pa"
malaise = "M1"
session = "S1"

# (sjsdmV = '0.1.2.9000') # package version
(sjsdmV <- packageVersion('sjSDM')[1]) # package version

(outputidxstatstabulatefolder <- glue("outputs_minimap2_{minimaprundate}_{samtoolsfilter}_{samtoolsqual}_kelpie{kelpierundate}_{primer}_vsearch97"))

# path to outputs from 1_sjsdm_model_20201024.Rmd
(pathtosjsdm <- here("..", "..", "Kelpie_maps", outputidxstatstabulatefolder, "crossvalidation_outputs"))

(cvfolder <- glue("results_{date.cross.validation}_{minocc}minocc_{envvar}_{abund}_loocv"))

(datafolder <- glue("data_{date.cross.validation}_{minocc}minocc_{envvar}"))

(otudata <- glue("otu.{abund}.csv"))

envdata <- "scale.env.csv"

# old
# # path to outputs from 1_sjsdm_model_20201024.Rmd
# pathtosjsdm <- here("..", "..", "sjSDM", "R-git", "results", "crossvalidation")
# 
# (cvfolder <- glue("results_{date.cross.validation}_{minocc}minocc_{envvar}_{abund}_loocv"))
# 
# (datafolder <- glue("data_{date.cross.validation}_{minocc}minocc_{envvar}"))
# 
# (otudata <- glue("otu.{abund}.csv"))
# 
# envdata <- "scale.env.csv"
```

```{r load model outputs and datasets}
result <- readRDS(file.path(pathtosjsdm, cvfolder,
                    glue('sjsdm_result_HJA_{date.cross.validation}.RDS')))
names(result)

model <- readRDS(file.path(pathtosjsdm, cvfolder,
                    glue('sjsdm_model_HJA_{date.cross.validation}.RDS')))
names(model)
model["cl"]
model["settings"]

best <- readRDS(file.path(pathtosjsdm, cvfolder,
    glue('sjsdm_tune_results_HJA_{date.cross.validation}_bestonly.RDS')))

summary.p <- readRDS(file.path(pathtosjsdm, cvfolder,
    glue("sjsdm_summary.p_HJA_{date.cross.validation}.RDS")))

imp <- readRDS(file.path(pathtosjsdm, cvfolder,
    glue('sjsdm_imp_HJA_{date.cross.validation}.RDS')))

otu <- read_csv(file.path(pathtosjsdm, cvfolder, datafolder, otudata))

env <- read_csv(file.path(pathtosjsdm, cvfolder, datafolder, envdata))
```

```{r graph importance ternary plot and barchart}
impdf <- tibble(OTU = imp$names, as_tibble(imp$res$total)) %>% 
    separate(OTU, 
             into = c("OTU", "empty", "class", "order", "family",
                      "genus", "epithet", "BOLD", "BOLDID",
                      "size"), remove = FALSE, sep = "_") %>% 
    select(-empty, -BOLD) %>% 
    relocate(env, spatial, biotic) 

pdf(here(pathtosjsdm, cvfolder, glue('importance_ternary_sjSDM_{malaise}_{session}_{abund}_min{minocc}_{envvar}_{date.cross.validation}.pdf')), width=8, height=5)

titleText=paste('Variation partition, ', nrow(data), ' OTUs', sep='')
# ind = 5 is Class:  Insecta, Arachnida
# ind = 6 is Order
vp.plot(data = impdf, x1 = 1, x2 = 3, ind = 6,  
        textM=titleText, textL="Order")
	
dev.off()

# barchart ranges of env, biotic, spatial
a = data.frame(range=names(summary(cut(impdf$biotic, breaks = seq(0,1,length.out = 12)))), env=summary(cut(impdf$env, breaks = seq(0,1,length.out = 12))), spatial=summary(cut(impdf$spatial, breaks = seq(0,1,length.out = 12))), biotic=summary(cut(impdf$biotic, breaks = seq(0,1,length.out = 12))), row.names=NULL)

a = melt(a, id="range", measure.vars=names(a)[2:4])
	
p1 <- ggplot(a, aes(range, value, fill = variable)) + 
    geom_bar(stat="identity", position = "dodge") +
    scale_fill_brewer(palette = "Set1") #+ ggtitle(legendText)

ggsave(here(pathtosjsdm, cvfolder,
        glue("importance_barchart_{date.cross.validation}_{abund}.pdf")), 
       width = 11.7, 
       height = 8.3, 
       device = "pdf",
       p1)
```

```{r graph correlation heatmap}
dim(result$sigma)# cov <- result$sigma 
# [1] 268 268

co.env.spp <- cov2cor(result$sigma)
rownames(co.env.spp) <- impdf$BOLDID
colnames(co.env.spp) <- impdf$BOLDID
# rownames(co.env.spp) <- 1:dim(result$sigma)[1]   # spp.names
# colnames(co.env.spp) <- 1:dim(result$sigma)[1]   # spp.names
	
range(co.env.spp)
	
cut.t = cut(co.env.spp, breaks = seq(-1,1, length.out = 12))
summary(cut.t)
rm(cut.t)

# plot species correlation heatmap
pdf(here(pathtosjsdm, cvfolder, glue("correlation_heatmap_{date.cross.validation}.pdf")), 
    width = 12, height = 12)
	
ggcorrplot(co.env.spp, hc.order = T, outline.color = "white", 
    insig = "blank", sig.level = 0.05, lab_size = 1, 
    show.legend = T, tl.cex = 4, 
    title=glue("sjSDM version: {sjsdmV}, crossval {date.cross.validation}, coef-alpha: {best[['alpha_coef']]}, lambda: {best[['lambda_coef']]}")
    )

dev.off()
```

```{r circular covariance plot by sorted species abundances}
otu.tbl <-  otu
number <- 10
	
sigma = re_scale(result$sigma)[order(apply(otu.tbl, 2, sum)),
                               order(apply(otu.tbl, 2, sum))]

pdf(here(pathtosjsdm, cvfolder, glue("circular_correlation_plot_sorted_by_species_abundance_{date.cross.validation}.pdf")), width = 8, height = 8)

version.text <- glue("sjSDM version: {sjsdmV}, crossval {date.cross.validation}, coef-alpha: {best[['alpha_coef']]}, lambda: {best[['lambda_coef']]}")

(otu.text <- glue("sum of OTU {abund}"))
	
cov.circle(version.text = version.text, otu.text = otu.text, 
           sigma = sigma, otu.tbl = otu.tbl, result = result)
	
dev.off()
```

```{r circular covariance plot sorted by env covariates}
beta = as.matrix(data.frame(result$beta$env, result$beta$spatial)[,2:29])

effects = apply(beta, 1, function(o) sum(abs(o)))

str(effects)
	
n = ncol(result$sigma)# number of otus

max_effects = apply(beta ,1, function(e) which.max(abs(e)))
	
effect_comb = data.frame(cbind(max_effects, 
                               sapply(1:n, function(i) beta[i, max_effects[i]] )))

str(effect_comb)

# variables index & value which has biggest coefficient

version.text <- glue("sjSDM version: {sjsdmV}, crossval {date.cross.validation}, coef-alpha: {best[['alpha_coef']]}, lambda: {best[['lambda_coef']]}")

(otu.text <- glue("sum of OTU {abund}"))

# need to check if order of env covariates is correct
evnames <- names(env)
evnames

pdf(here(pathtosjsdm, cvfolder, glue("circular_correlation_plot_sorted_by_env_covariates_{date.cross.validation}.pdf")), width = 8, height = 8)

cov.circle.env(version.text = version.text, evnames = evnames,
               otu.text = otu.text, result = result, 
               effect_comb = effect_comb, otu.tbl = otu.tbl)
	
dev.off()
	
```

```{r env coefficients barchart}
# set variables
minsize <- 10
maxsize <- 200000L # default 200000 to include all large OTUs
taxon <- "all" # "all" is the default

# function(summary.p, result, minsize, maxsize=200000, taxon="all")
p1 <- coef.figure(summary.p, result, minsize, maxsize, taxon)
# p1

ggsave(here(pathtosjsdm, cvfolder,
    glue("coef_figure_min{minsize}_max{maxsize}_{taxon}-taxa_{date.cross.validation}.pdf")), 
       width = 15, 
       height = 24, 
       device = "pdf",
       p1
    )

```

```{r histograms of env coefficients and otu sizes}
effect <- data.frame(summary.p$coefmat)
effect$rownames <- rownames(effect)
effect <- separate(
    data = effect,
    col = rownames,
    into = c("species", "coef"),
    sep = " "
    ) %>% 
    filter(coef != "(Intercept)")

effect <- effect %>%
    tidyr::separate(species, into = c("OTU", "taxon"), sep = "__", remove = TRUE) %>%
    tidyr::separate(taxon, into = c("taxon", "size"), sep = "_size.", remove = TRUE) %>%
    # mutate(taxon = str_replace(taxon, "_sp", "_genus_sp")) %>%
    tidyr::separate(taxon, into = c("class", "order", "family", "genus", "species", "BOLD", "BOLDID")) %>%
    unite(OTU, c("OTU", "class", "order", "family", "genus", "species"), sep = "_", ) %>%
    select(-BOLD, -BOLDID) %>%
    mutate(size = as.numeric(size)) %>%
    rename(species = OTU)

effecthist <- effect %>% distinct(species, .keep_all = TRUE)
hist(effecthist$Estimate)
hist(effecthist$size, breaks = 5000, xlim = c(0, 10000))
```

