---
title: "ecocopula"
author: "Douglas Yu"
date: "04/10/2020"
output: html_document
---

```{r setup, eval=TRUE, include=TRUE}
# script-specific libraries
suppressPackageStartupMessages({
  library(stringdist)
  library(vegan)
  library(mvabund)
  library(ordinal)
  library(ecoCopula)
  library(corrplot)
  library(ggrepel)
  library(RColorBrewer)
  library(tidygraph)
  library(ggraph)
})

# general-use packages
suppressPackageStartupMessages({
  library(tidyverse) # loads a bunch of packages (see below)
  library(here)
  library(fs)
  library(glue)
  library(readxl)
  library(cowplot)
  library(lubridate)
  library(patchwork)
  library(broom)
  library(ggeffects)
  library(viridis)
  library(arsenal) # for tableby()
  library(waldo) # for compare()
  library(sjmisc) # for rotate_df()
  library(envDocument)
  library(inspectdf)
  library(conflicted)
  library(knitr)
  library(beepr)
  library(pivottabler)
  library(furrr)
  library(scales)
  library(janitor)
  library(tictoc)
})

conflict_prefer("mutate", "dplyr", quiet = TRUE)
conflict_prefer("select", "dplyr", quiet = TRUE)
conflict_prefer("summarise", "dplyr", quiet = TRUE)
conflict_prefer("filter", "dplyr", quiet = TRUE)
conflict_prefer("first", "dplyr", quiet = TRUE)
conflict_prefer("here", "here", quiet = TRUE)
conflict_prefer("separate", "tidyr", quiet = TRUE)
conflict_prefer("unite", "tidyr", quiet = TRUE)
conflict_prefer("trim", "sjmisc", quiet=TRUE)
conflict_prefer("rescale", "scales", quiet=TRUE)
conflict_prefer("rescale", "discard", quiet=TRUE)
conflict_prefer("intersect", "dplyr", quiet = TRUE)
conflict_prefer("setdiff", "dplyr", quiet = TRUE) # w/out this, R crashes
conflict_prefer("to_factor", "sjmisc", quiet = TRUE)
conflict_prefer("trim", "glue", quiet = TRUE)
conflict_prefer("discard", "purrr", quiet = TRUE)

# Real numbers, not scientific notation.
options(scipen = 999)
```

```{r working directory}
here() # should be: "/Users/Negorashi2011/Dropbox/Working_docs/Luo_Mingjie_Oregon/HJA_analyses_Kelpie/HJA_scripts/09_ecoCopula_analyses"
```

read in the uncorr otuenv file.  uncorr means no log(FSL) correction, no QP transformation. This gives us the raw read numbers for ecoCopula analysis
```{r read data}
samtoolsfilter <- "F2308" # F2308 filter only
samtoolsqual <- "q48"
minimaprundate <- 20200929
kelpierundate <- 20200927
primer <- "BF3BR2"

(outputidxstatstabulatefolder <- glue("outputs_minimap2_{minimaprundate}_{samtoolsfilter}_{samtoolsqual}_kelpie{kelpierundate}_{primer}_vsearch97"))

# read uncorr sample x species table
otuenv <- read_csv(here("..", "..", "Kelpie_maps", 
                        outputidxstatstabulatefolder, glue("sample_by_species_table_{samtoolsfilter}_minimap2_{minimaprundate}_kelpie{kelpierundate}_uncorr.csv")))
```

```{r subset otuenv by trap and period}
# M1S1
trap <- "M1"
period <- "S1"
otuenv <- otuenv %>% 
  filter(trap == trap[[1]] & period == period[[1]]) 
```

```{r create and scale XY.csv}
XY.csv <- otuenv %>% 
    select(UTM_E, UTM_N) %>% 
    scale() %>% 
    as_tibble()
```

```{r create otu.ab.csv, otu.qp.csv, and otu.pa.csv}
# keep OTUs with >=5 incidences
# original read number abundance
minocc <- 5 # set to high number (e.g. 20) for testing
otu.ab.csv <- otuenv %>% select(lysis_ratio, contains("__")) %>% 
  mutate(across(contains("__"), ~ .x / lysis_ratio)) %>% 
  select(-lysis_ratio) %>% 
  round()
otu.ab.csv <- otu.ab.csv[ , specnumber(otu.ab.csv, MARGIN = 2) >= minocc] 

# Before quasiprobability transformation, check that all OTUs have at least one cell with 0 reads. Otherwise, the lowest read value will be QP-scaled to 0
OTUs_mins <- otu.ab.csv %>% 
  summarise(across(contains("__"), min))
rowSums(OTUs_mins) == 0 # # If TRUE, can go ahead with QP transformation because all cols have a 0 value
# log(FSL) correction and scale to quasiprobability
otu.qp.csv <- otu.ab.csv %>% 
  mutate(across(contains("__"), ~ .x /(otuenv$COISpike_sum))) %>% 
  mutate(across(contains("__"), ~ log(.x + 0.001))) %>% 
  mutate(across(contains("__"), ~ rescale(.x))) # {scales}
max(otu.qp.csv) == 1 # should be TRUE

# convert to presence/absence data
otu.pa.csv <- otu.ab.csv
otu.pa.csv[otu.pa.csv > 0] <- 1
min(colSums(otu.pa.csv)) == minocc # should be TRUE
```

otu.ab.csv has original read numbers corrected only for lysis buffer ratio
otu.qp.csv has been corrected for COIspike number, logged, and then qp scaled
otu.pa.csv has been converted to presence/absence

We use otu.ab.csv + offset(log(COISpike_sum)) instead of otu.qp.csv.  Maybe comment out the code that creates otu.qp.csv?

```{r create env.csv}
# env covariates
otuenv %>% 
    select(!contains("__"), -UTM_E, -UTM_N, -starts_with("nor")) %>% 
    names(.)
#   [1] "SiteName"           "trap"               "period"            
#   [4] "lysis_ratio"        "COISpike_sum"       "be10"              
#   [7] "tri"                "slope"              "aspect"            
#  [10] "Nss"                "Ess"                "twi"               
#  [13] "ht"                 "ht.r250"            "ht.r500"           
#  [16] "ht.r1k"             "cov2_4"             "cov2_4.r250"       
#  [19] "cov2_4.r500"        "cov2_4.r1k"         "cov4_16"           
#  [22] "cov4_16.r250"       "cov4_16.r500"       "cov4_16.r1k"       
#  [25] "be500"              "mTopo"              "cut.r1k.pt"        
#  [28] "clearcut"           "insideHJA"          "oldGrowthIndex"    
#  [31] "elevation_f"        "canopyHeight_f"     "minT_annual"       
#  [34] "maxT_annual"        "precipitation_mm"   "distToRoad_m"      
#  [37] "distToStream_m"     "YrsSinceDist"       "B1_20180717"       
#  [40] "B2_20180717"        "B3_20180717"        "B4_20180717"       
#  [43] "B5_20180717"        "B6_20180717"        "B7_20180717"       
#  [46] "B10_20180717"       "B11_20180717"       "NDVI_20180717"     
#  [49] "EVI_20180717"       "B_20180717"         "G_20180717"        
#  [52] "W_20180717"         "B1_20180726"        "B2_20180726"       
#  [55] "B3_20180726"        "B4_20180726"        "B5_20180726"       
#  [58] "B6_20180726"        "B7_20180726"        "B10_20180726"      
#  [61] "B11_20180726"       "NDVI_20180726"      "EVI_20180726"      
#  [64] "B_20180726"         "G_20180726"         "W_20180726"        
#  [67] "B1_20180802"        "B2_20180802"        "B3_20180802"       
#  [70] "B4_20180802"        "B5_20180802"        "B6_20180802"       
#  [73] "B7_20180802"        "B10_20180802"       "B11_20180802"      
#  [76] "NDVI_20180802"      "EVI_20180802"       "B_20180802"        
#  [79] "G_20180802"         "W_20180802"         "B1_20180818"       
#  [82] "B2_20180818"        "B3_20180818"        "B4_20180818"       
#  [85] "B5_20180818"        "B6_20180818"        "B7_20180818"       
#  [88] "B10_20180818"       "B11_20180818"       "NDVI_20180818"     
#  [91] "EVI_20180818"       "B_20180818"         "G_20180818"        
#  [94] "W_20180818"         "mean.NDVI"          "mean.EVI"          
#  [97] "mean.bright"        "mean.green"         "mean.wet"          
# [100] "mean.NDVI.scale"    "mean.EVI.scale"     "mean.green.scale"  
# [103] "mean.bright.scale"  "mean.wet.scale"     "l_Cover_2m_max"    
# [106] "l_Cover_2m_max_all" "l_Cover_2m_4m"      "l_Cover_2m_4m_all" 
# [109] "l_Cover_4m_16m"     "l_p25"              "l_p25_all"         
# [112] "l_p95"              "l_p95_all"          "l_rumple"      

# select subset of covariates chosen by Christian
# remove OTUs, XY, and normalised NDVI and EVI
# log some of the distance and cover variates
env.csv <- otuenv %>% 
    select(!contains("__"), -UTM_E, -UTM_N, -starts_with("nor")) %>%
    mutate(lg_DistStream = log(distToStream_m + 0.001)) %>% 
    mutate(lg_DistRoad = log(distToRoad_m + 0.001)) %>% 
    mutate(lg_YrsDisturb = log(YrsSinceDist + 0.001)) %>% 
    mutate(lg_cover2m_max = log(l_Cover_2m_max + 0.001)) %>% 
    mutate(lg_cover2m_4m = log(l_Cover_2m_4m + 0.001)) %>%
    mutate(lg_cover4m_16m = log(l_Cover_4m_16m + 0.001)) %>%   
    select(SiteName, trap, period, lysis_ratio, COISpike_sum, clearcut, insideHJA, oldGrowthIndex, elevation_f, canopyHeight_f, precipitation_mm, be10:cut.r1k.pt, mean.NDVI, mean.EVI, mean.green, mean.wet, l_p25, l_rumple, contains("20180717"), lg_DistStream, lg_DistRoad, lg_YrsDisturb, lg_cover2m_max, lg_cover2m_4m, lg_cover4m_16m) %>%
  mutate(across(where(is.numeric) & !lysis_ratio & !COISpike_sum, scale)) %>%
  mutate(offset = (COISpike_sum / lysis_ratio)) %>% 
  relocate(offset, .after = COISpike_sum)
```

the smaller the lysis_ratio, the larger the original sample, and i just increase the number of reads by the lysis_ratio (e.g. if the lysis_ratio is 0.5, the numbers of reads in that row are increased by 2X)
COISpikesum/lysis_ratio = offset is proportional to the observation 'effort' during the metabarcoding pipeline
I thus use offset(log(offset))

Functions to plot ecoCopula output (cord())
```{r process cord output}
site_res_fxn <- function(cordord, env, xy) {
  data.frame(cordord$scores, env, xy)
  }

sp_res_fxn <- function(cordord, abund) {
  data.frame(cordord$loadings, 
             species = colnames(get(glue("otu.{abund}.csv")))) %>%
  separate(species, into = c("OTU", "empty", "class", "order", "family",
                               "genus", "epithet", "BOLD", "BOLDID",
                               "size"), remove = FALSE, sep = "_") %>%
  select(-empty)
  }

plot_factors <- function(alphanum, abund, model) {
  alpha <- alphanum
  ggplot() + 
  geom_segment(aes(x = 0, y = 0, 
                   xend = Factor1 * alpha * 0.95, 
                   yend = Factor2 * alpha * 0.95), 
               data = sp_res, 
               size = .1) +
  geom_point(aes(x = Factor1, y = Factor2, 
                 color = be500, 
                 size = exp(lg_YrsDisturb), 
                 shape = as.factor(insideHJA)
                 ), 
             data = site_res) + 
  geom_text_repel(aes(x = Factor1, y = Factor2, 
                      label = env.csv$SiteName), 
                  data = site_res, 
                  size = 2,
                  max.overlaps = 15) +
  # geom_label_repel(aes(x = Factor1 * alpha, y = Factor2 * alpha, label = species), data = sp_res, size = 2.5) +
  scale_color_gradientn(colours = brewer.pal(n = 10, name = "RdYlBu")) +
  theme_classic() + # or "PuOr" or "RdYlBu"
  labs(title = glue("HJA, ecoCopula ordination, {abund}"), 
       subtitle = model) +
  xlab("Factor 1") +
  ylab("Factor 2")
  }

plot_xy_factor1 <- function(abund, model) {
  ggplot() + 
  geom_point(aes(x = UTM_E, y = UTM_N, 
                 color = Factor1, 
                 size = exp(lg_YrsDisturb), # YrsSinceDist, l_rumple
                 shape = as.factor(insideHJA)
                 ), 
             data = site_res) +
  geom_text_repel(aes(x = UTM_E, y = UTM_N, 
                      label = env.csv$SiteName), 
                  data = site_res, 
                  size = 2,
                  max.overlaps = 15) +
  # geom_label_repel(aes(x = Factor1 * alpha, y = Factor2 * alpha, label = species), data = sp_res, size = 2.5) +
  scale_color_gradientn(colours = brewer.pal(n = 10, name = "PuOr")) +
  theme_classic() + # or "PuOr" or "RdYlBu"
  labs(title = glue("HJA, XY plot, Factor1, {abund}"),
       subtitle = model) +
  xlab("UTM_E") +
  ylab("UTM_N")
}

plot_xy_factor2 <- function(abund, model) {
  ggplot() + 
  geom_point(aes(x = UTM_E, y = UTM_N, 
                 color = Factor2, 
                 size = exp(lg_YrsDisturb), # YrsSinceDist, l_rumple
                 shape = as.factor(insideHJA)
                 ), 
             data = site_res) +
  geom_text_repel(aes(x = UTM_E, y = UTM_N, 
                      label = env.csv$SiteName), 
                  data = site_res, 
                  size = 2,
                  max.overlaps = 15) +
  # geom_label_repel(aes(x = Factor1 * alpha, y = Factor2 * alpha, label = species), data = sp_res, size = 2.5) +
  scale_color_gradientn(colours = brewer.pal(n = 10, name = "PuOr")) +
  theme_classic() + # or "PuOr" or "RdYlBu"
  labs(title = glue("HJA, XY plot, Factor2, {abund}"),
       subtitle = model) +
  xlab("UTM_E") +
  ylab("UTM_N")
}

plot_corrplot <- function(abund, model) {
  site_res_num <- site_res %>% 
    select(-SiteName, -trap, -period, -lysis_ratio, -COISpike_sum,
           -clearcut, -insideHJA)
  corrplot(cor(site_res_num), 
           method = "ellipse", 
           type = "lower", 
           title = glue("{abund}, {model}"),
           mar=c(0,0,2,0)
           )
  # # save corrplot
  pdf(file = here("..", "..", "Kelpie_maps",
                outputidxstatstabulatefolder,
                glue("corrplot_{trap}{period}_minimap2_{minimaprundate}_kelpie{kelpierundate}_{abund}.pdf")
                ),
      width = 11.7,
      height = 8.3
    )
  corrplot(cor(site_res_num),
           method = "ellipse",
           type = "lower",
           title = glue("{abund}, {model}"),
           mar=c(0,0,2,0)
           )
  dev.off()
}
```


mvabund models, mod0 and mod1
```{r manyglm pa}
abund <- "pa"
glue("otu.{abund}.csv")

otu.ord <- mvabund(get(glue("otu.{abund}.csv")))
is.mvabund(otu.ord)

# with no env covariates
otu.mod0 <- manyglm(otu.ord ~ 1,
                    family = binomial("cloglog"), # family = "binomial"
                    data = env.csv
                    )
plot(otu.mod0) # chk residuals

otu.mod0.ord <- cord(otu.mod0)

# with elevation
otu.mod1 <- manyglm(otu.ord ~ elevation_f, 
                    family = binomial("cloglog"), # family = "binomial"
                    data = env.csv
                    )
plot(otu.mod1) # chk residuals

otu.mod1.ord <- cord(otu.mod1)

# make plots

# choose model
# mod0: no covariates
cordord <- "otu.mod0.ord"; model <- "otu.ord ~ 1"
# mod1: elevation_f 
cordord <- "otu.mod1.ord"; model <- "otu.ord ~ elevation_f"

site_res <- site_res_fxn(cordord = get(cordord), env.csv, XY.csv)
sp_res <- sp_res_fxn(cordord =  get(cordord), abund = abund)
p0 <- plot_factors(alphanum = 4, 
                   abund = abund, 
                   model = model
                   )
p1 <- plot_xy_factor1(abund = abund, model = model)
p2 <- plot_xy_factor2(abund = abund, model = model)

p0
p1 | p2 # if "Viewport has zero dimension(s)" error, enlarge plot window

# make plots and save

plot_corrplot(abund = abund, model = model)

ggsave(here("..", "..", "Kelpie_maps", 
            outputidxstatstabulatefolder,
            glue("ecoCopula_{trap}{period}_minimap2_{minimaprundate}_kelpie{kelpierundate}_{abund}.pdf")
            ), 
       width = 11.7, 
       height = 8.3, 
       device = "pdf",
       p0)

ggsave(here("..", "..", "Kelpie_maps", 
            outputidxstatstabulatefolder,
            glue("ecoCopulaXY_{trap}{period}_minimap2_{minimaprundate}_kelpie{kelpierundate}_{abund}.pdf")
            ), 
       width = 11.7, 
       height = 8.3, 
       device = "pdf",
       p1 | p2)
```

manyglm models, mod2 and mod3. ab means original read counts, with log(offset) to control for lysis buffer and COI spike
```{r manyglm ab}
abund <- "ab"
glue("otu.{abund}.csv")
otu.ord <- mvabund(get(glue("otu.{abund}.csv")))
is.mvabund(otu.ord)

otu.mod2 <- manyglm(otu.ord ~ 1 + offset(log(COISpike_sum)), 
                    family = "negative.binomial", 
                    data = env.csv
                    )
plot(otu.mod2) # chk residuals
otu.mod2.ord <- cord(otu.mod2)

# with be500
otu.mod3 <- manyglm(otu.ord ~ be500 + offset(log(COISpike_sum)), 
                    family = "negative.binomial", 
                    data = env.csv
                    )
plot(otu.mod3) # chk residuals
otu.mod3.ord <- cord(otu.mod3)

# with all variables that look correlated with Factor 1
otu.mod6 <- manyglm(otu.ord ~ be500 + Nss + B1_20180717 + 
                      B2_20180717 + B3_20180717 + B4_20180717 +
                      B6_20180717 + B7_20180717 + NDVI_20180717 +
                      lg_DistRoad + mean.wet + l_p25 +
                      offset(log(COISpike_sum)), 
                    family = "negative.binomial", 
                    data = env.csv,
                    show.time = "all"
                    )
plot(otu.mod6) # chk residuals
otu.mod6.ord <- cord(otu.mod6)
# anova(otu.mod6, nBoot = 100) 

# make plots
# mod2: no covariates
cordord <- "otu.mod2.ord"; model <- "otu.ord ~ 1" 
# OR
# mod3: elevation_f 
cordord <- "otu.mod3.ord"; model <- "otu.ord ~ be500"
# OR
# mod6: elevation_f 
cordord <- "otu.mod6.ord"; model <- "otu.ord ~ 12 covariates"


site_res <- site_res_fxn(cordord = get(cordord), env.csv, XY.csv)
sp_res <- sp_res_fxn(cordord =  get(cordord), abund = abund)
p0 <- plot_factors(alphanum = 4, 
                   abund = abund, 
                   model = model
                   )
p1 <- plot_xy_factor1(abund = abund, model = model)
p2 <- plot_xy_factor2(abund = abund, model = model)

p0
p1 | p2 # if "Viewport has zero dimension(s)" error, enlarge plot window

# make plots and save

plot_corrplot(abund = abund, model = model) # need to update site_res with active model for this to work

ggsave(filename = here("..", "..", "Kelpie_maps", 
                outputidxstatstabulatefolder,
                glue("ecoCopula_{trap}{period}_minimap2_{minimaprundate}_kelpie{kelpierundate}_{abund}.pdf")
                ), 
       width = 11.7, 
       height = 8.3, 
       device = "pdf",
       p0)

ggsave(filename = here("..", "..", "Kelpie_maps", 
                outputidxstatstabulatefolder,
                glue("ecoCopulaXY_{trap}{period}_minimap2_{minimaprundate}_kelpie{kelpierundate}_{abund}.pdf")
                ), 
       width = 11.7, 
       height = 8.3, 
       device = "pdf",
       p1 | p2)
```

Manyany models, mod4 and mod5
```{r Manyany qp}
source(here("1.1_Manyany.R"))

abund <- "qp"
glue("otu.{abund}.csv")

otu.ord <- get(glue("otu.{abund}.csv")) %>% 
  mutate(across(everything(), factor))

# Now fit a model using the clm function:
otu.mod4 <- Manyany("clm", otu.ord, abund ~ 1, data = env.csv)

plot(otu.mod4) # chk residuals

otu.mod4.ord <- cord(otu.mod4)

# with elevation
otu.mod5 <- Manyany("clm", otu.ord, abund ~ elevation_f, data = env.csv)

plot(otu.mod5) # chk residuals

otu.mod5.ord <- cord(otu.mod5)

# make plots
# mod4: no covariates
cordord <- "otu.mod4.ord"; model <- "otu.ord ~ 1"
# OR
# mod5: elevation_f 
cordord <- "otu.mod5.ord"; model <- "otu.ord ~ elevation_f"

site_res <- site_res_fxn(cordord = get(cordord), env.csv, XY.csv)
sp_res <- sp_res_fxn(cordord =  get(cordord), abund = abund)
p0 <- plot_factors(alphanum = 4, 
                   abund = abund, 
                   model = model
                   )
p1 <- plot_xy_factor1(abund = abund, model = model)
p2 <- plot_xy_factor2(abund = abund, model = model)

p0
p1 | p2 # if "Viewport has zero dimension(s)" error, enlarge plot window

# make plots and save

plot_corrplot(abund = abund, model = model)

ggsave(filename = here("..", "..", "Kelpie_maps", 
                outputidxstatstabulatefolder,
                glue("ecoCopula_{trap}{period}_minimap2_{minimaprundate}_kelpie{kelpierundate}_{abund}.pdf")
                ), 
       width = 11.7, 
       height = 8.3, 
       device = "pdf",
       p0)

ggsave(filename = here("..", "..", "Kelpie_maps", 
                outputidxstatstabulatefolder,
                glue("ecoCopulaXY_{trap}{period}_minimap2_{minimaprundate}_kelpie{kelpierundate}_{abund}.pdf")
                ), 
       width = 11.7, 
       height = 8.3, 
       device = "pdf",
       p1 | p2)
```



manyglm models, mod4.1 and mod5.1. *No offset term*.  ab means read counts are kept, with log(offset) to control for lysis buffer and COI spike
```{r manyglm ab}
abund <- "ab"
glue("otu.{abund}.csv")
otu.ord <- mvabund(get(glue("otu.{abund}.csv")))
is.mvabund(otu.ord)

# no log(offset) term
otu.mod4.1 <- manyglm(otu.ord ~ 1, 
                    family = "negative.binomial", 
                    data = env.csv,
                    composition = TRUE
                    )
plot(otu.mod4.1) # chk residuals
otu.mod4.1.ord <- cord(otu.mod4.1)

# with elevation. no log(offset) term
otu.mod5.1 <- manyglm(otu.ord ~ elevation_f, 
                    family = "negative.binomial", 
                    data = env.csv,
                    composition = TRUE
                    )
plot(otu.mod5.1) # chk residuals
otu.mod5.1.ord <- cord(otu.mod5.1)

# make plots
# mod4.1: no covariates, no offset term
# mod5.1: elevation_f, no offset term
cordord <- "otu.mod4.1.ord" # "otu.mod4.1.ord" # "otu.mod5.1.ord"
model <- "otu.ord ~ 1, no offset" # "otu.ord ~ 1, no offset" # "otu.ord ~ elevation_f, no offset"
site_res <- site_res_fxn(cordord = get(cordord), env.csv, XY.csv)
sp_res <- sp_res_fxn(cordord =  get(cordord), abund = abund)
p0 <- plot_factors(alphanum = 4, 
                   abund = abund, 
                   model = model
                   )
p1 <- plot_xy_factor1(abund = abund, model = model)
p2 <- plot_xy_factor2(abund = abund, model = model)

p0
p1 | p2 # if "Viewport has zero dimension(s)" error, enlarge plot window

# make plots and save

plot_corrplot(abund = abund, model = model)

ggsave(filename = here("..", "..", "Kelpie_maps", 
                outputidxstatstabulatefolder,
                glue("ecoCopula_{trap}{period}_minimap2_{minimaprundate}_kelpie{kelpierundate}_{abund}.pdf")
                ), 
       width = 11.7, 
       height = 8.3, 
       device = "pdf",
       p0)

ggsave(filename = here("..", "..", "Kelpie_maps", 
                outputidxstatstabulatefolder,
                glue("ecoCopulaXY_{trap}{period}_minimap2_{minimaprundate}_kelpie{kelpierundate}_{abund}.pdf")
                ), 
       width = 11.7, 
       height = 8.3, 
       device = "pdf",
       p1 | p2)
```


```{r}
cordord <- "otu.mod8.ord"; model <- "otu.ord ~ lg_YrsDisturb + elevation_f * insideHJA"
env <- otuenv %>% 
  select(!contains("__"), -UTM_E, -UTM_N, -starts_with("nor")) %>% 
  select(!contains(c("20180726", "20180818", "20180802"))) %>% 
  select(!ends_with("_all"))
  
site_res <- site_res_fxn(cordord = get(cordord), env, XY.csv)
sp_res <- sp_res_fxn(cordord =  get(cordord), abund = abund)

plot_corrplot(abund = abund, model = model)
```


From the Gordana Popovic tutorial:  "We can do this using the `mvabund` package, with function `manyany` as well as `clm` from the `ordinal` package. Sadly, R version 4.0.0 has broken `manyany`, and it has not been fixed yet, so I have a substitute called `Manyany` in the `1.1_Manyany.R` file which we will need to sourced in the code below. "

Manyany<-function (fn, yMat, formula, data, family = "negative.binomial", 
          composition = FALSE, block = NULL, get.what = "details", 
          var.power = NA, na.action = "na.exclude", ...) 
          

<details>
<summary>Reproducibility receipt</summary>
```{r reproducibility receipt}
# datetime
Sys.time()

# repository
git2r::repository(here::here())

envDocument::env_doc("table", git = FALSE)
```
</details>


Deprecate
```{r plot ecoCopula ordination, include=FALSE, eval=FALSE}
alpha <- 4
p0 <- ggplot() + 
  geom_segment(aes(x = 0, y = 0, 
                   xend = Factor1 * alpha * 0.95, 
                   yend = Factor2 * alpha * 0.95), 
               data = sp_res, 
               size = .1) +
  geom_point(aes(x = Factor1, y = Factor2, 
                 color = elevation_f, 
                 size = exp(lg_YrsDisturb), 
                 shape = as.factor(insideHJA)
                 ), 
             data = site_res) + 
  geom_text_repel(aes(x = Factor1, y = Factor2, 
                      label = env.csv$SiteName), 
                  data = site_res, 
                  size = 2) +
  # geom_label_repel(aes(x = Factor1 * alpha, y = Factor2 * alpha, label = species), data = sp_res, size = 2.5) +
  scale_color_gradientn(colours = brewer.pal(n = 10, name = "RdYlBu")) +
  theme_classic() + # or "PuOr" or "RdYlBu"
  labs(title = glue("HJA, ecoCopula ordination, {abund}"))

p0
```

```{r plot by XY position, include=FALSE, eval=FALSE}
p1 <- ggplot() + 
  geom_point(aes(x = UTM_E, y = UTM_N, 
                 color = Factor1, 
                 size = exp(lg_YrsDisturb), # YrsSinceDist, l_rumple
                 shape = as.factor(insideHJA)
                 ), 
             data = site_res) +
  geom_text_repel(aes(x = UTM_E, y = UTM_N, 
                      label = env.csv$SiteName), 
                  data = site_res, 
                  size = 2) +
  # geom_label_repel(aes(x = Factor1 * alpha, y = Factor2 * alpha, label = species), data = sp_res, size = 2.5) +
  scale_color_gradientn(colours = brewer.pal(n = 10, name = "PuOr")) +
  theme_classic() + # or "PuOr" or "RdYlBu"
  labs(title = glue("HJA, XY plot, Factor1, {abund}"))

p2 <- ggplot() + 
  geom_point(aes(x = UTM_E, y = UTM_N, 
                 color = Factor2, 
                 size = exp(lg_YrsDisturb), # YrsSinceDist, l_rumple
                 shape = as.factor(insideHJA)
                 ), 
             data = site_res) +
  geom_text_repel(aes(x = UTM_E, y = UTM_N, 
                      label = env.csv$SiteName), 
                  data = site_res, 
                  size = 2) +
  # geom_label_repel(aes(x = Factor1 * alpha, y = Factor2 * alpha, label = species), data = sp_res, size = 2.5) +
  scale_color_gradientn(colours = brewer.pal(n = 10, name = "PuOr")) +
  theme_classic() + # or "PuOr" or "RdYlBu"
  labs(title = glue("HJA, XY plot, Factor2, {abund}"))

p1 | p2
```
